---
title: "Educational Learning Analytics - Classification problem"
author: "Krisztina Gyalai"
date: "19/01/2022"
output: html_document
---

**Aim** test 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#load tidyverse
library(tidyverse)
```

## Data preparation and attribute selection 

Dataset consist of 7 files which fits into three categories:

1. Student Activities
2. Student Demographics 
3. Module Presentation 
  
The full list of attributes and their description available at the [Open Dataset Webpage](https://analyse.kmi.open.ac.uk/open_dataset)

There are three required files identified: 
1. studentInfo: All required background information of the student including the target feature: 
2. studentVle: Contain information of the student engagement with the study materials (number of clicks on website) treated as main predictor. 
3. studentRegistration: students with withdrawal date < 0 need to be eliminated from the study as they left course before its started they success/failure cannot be measured - would result noise to the model. 

```{r echo=TRUE}
#load datasets 
main <- read_csv("~/git/OU Learning Analytics/Raw Data/studentInfo.csv",show_col_types = FALSE)
reg <- read_csv("~/git/OU Learning Analytics/Raw Data/studentRegistration.csv",show_col_types = FALSE)
vle <- read_csv("~/git/OU Learning Analytics/Raw Data/studentVle.csv",show_col_types = FALSE)

#merge student registration into the main data set 
#unique identifier student id
main_reg <- merge(main, reg, by ="id_student")

```

According to the data source the student id is a unique identifier across the 7 data files. However, by merging the studentRegistration file to the studentInfo, the row numbers increases by `r 40801-32593` indicates introduced duplicates. 

Investigating duplicates: 
```{r echo=TRUE}
#find duplicates in merged df (in student id)
duplicates <- main_reg %>%
  group_by(id_student)%>%
  summarise(n=n())%>%
  filter(n>1)
#counts 
unique(duplicates$n)
```

In some extreme cases student number occurs 25 times within the merged df. 

```{r echo=TRUE}
#filter extreme cases
duplicates %>%
  filter(n >= 16)

```
```{r echo=TRUE, fig.length = 2}
#randomly select one student to see details - 584077 n = 25
main_reg %>%
  filter(id_student == 584077)%>%
  select(code_module.x,code_presentation.x,date_registration,date_unregistration)

```
The above example show where and while is the error introduced through merging. 
**student id** is unique identifier of the student, but **not a unique identifier** across the data set. 
As the above example show: 

  * It is possible for the student to enroll to the same module (code_module) more than once.  
  * It is possible for the student to enroll to the same module in different presentations (code_presentation) more than once. 
  
  * It is NOT possible to the student to enroll to the same module and same presentation more than once. 

Therefore to avoid to introduce duplicates and errors through merging, the merging must be based on ALL three attributes - the 3 required df need to be broken up per modules and per presentation before merging. 

```{r}
#AAA
#separate main file by module and presentation 
unique(main$code_module)
main_A <- main %>%
  filter(code_module == "AAA")
#2013J
unique(main_A$code_presentation)
A_2013J <- main_A%>%
  filter(code_presentation == "2013J")
#2014J
A_2014J <- main_A%>%
  filter(code_presentation == "2014J")

#registration 
#select only the columns that want to keep
reg_A <- reg %>%
  filter(code_module =="AAA")
reg_2013J <- reg_A %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,date_unregistration)
reg_2014J <- reg_A %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,date_unregistration)
  

#vle 
#group by student number
#select only the columns that want to keep 
vle_A <- vle %>%
  filter(code_module == "AAA")
vle_2013J <- vle_A %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
#2013J vle 5 row less than main and reg - check how much students unregistered before start date
reg_2013J%>%
  filter(date_unregistration <= 0)

#there are 7 student unregistered before module started but only 5 missing in vle 
#there are still some discrepancy but is greatly reduced  

vle_2014J <- vle_A %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merge main and registration 2013J
mr_A_2013J <- merge(A_2013J,reg_2013J,by="id_student")
#add vle
mrv_A_2013J <- merge(mr_A_2013J, vle_2013J, by="id_student")

#repeat for 2014J
mr_A_2014J <- merge(A_2014J, reg_2014J, by="id_student")
#add vle 
mrv_A_2014J <- merge(mr_A_2014J, vle_2014J, by="id_student")

#combine the two files together to archive the merged data frame for module AAA
AAA <- rbind(mrv_A_2013J, mrv_A_2014J)

```

Breaking up files per presentation and by module type then combine them back together is successfully eliminate the errors, therefore needs to be repeated to all seven presentations individually. 

```{r}
#BBB
main_B <- main %>%
  filter(code_module == "BBB")
#check how many presentations there
unique(main_B$code_presentation)
#there are four different files needed 
main_B_2013B <- main_B%>%
  filter(code_presentation == "2013B")
main_B_2013J <- main_B %>%
  filter(code_presentation =="2013J")
main_B_2014B <- main_B %>%
  filter(code_presentation == "2014B")
main_B_2014J <- main_B %>%
  filter(code_presentation == "2014J")

#break registration
reg_B <- reg %>%
  filter(code_module == "BBB")
reg_B_2013B <- reg_B%>%
  filter(code_presentation == "2013B")%>%
    select(id_student,date_unregistration)
reg_B_2013J <- reg_B %>%
  filter(code_presentation =="2013J")%>%
    select(id_student,date_unregistration)
reg_B_2014B <- reg_B %>%
  filter(code_presentation == "2014B")%>%
    select(id_student,date_unregistration)
reg_B_2014J <- reg_B %>%
  filter(code_presentation == "2014J")%>%
    select(id_student,date_unregistration)

#vle
vle_B <- vle %>%
  filter(code_module == "BBB")
vle_B_2013B <- vle_B %>%
  filter(code_presentation == "2013B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_B_2013J <- vle_B %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_B_2014B <- vle_B %>%
  filter(code_presentation == "2014B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_B_2014J <- vle_B %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merging 
#merge main and registration 2013B
mr_B_2013B <- merge(main_B_2013B,reg_B_2013B,by="id_student")
#add vle
mrv_B_2013B <- merge(mr_B_2013B, vle_B_2013B, by="id_student")
#B_2013J
mr_B_2013J <- merge(main_B_2013J,reg_B_2013J,by="id_student")
#add vle
mrv_B_2013J <- merge(mr_B_2013J, vle_B_2013J, by="id_student")
#2014_B
mr_B_2014B <- merge(main_B_2014B,reg_B_2014B,by="id_student")
#add vle
mrv_B_2014B <- merge(mr_B_2014B, vle_B_2014B, by="id_student")
#B_2014J
mr_B_2014J <- merge(main_B_2014J,reg_B_2014J,by="id_student")
#add vle
mrv_B_2014J <- merge(mr_B_2014J, vle_B_2014J, by="id_student")

BBB <- rbind(mrv_B_2013B,mrv_B_2013J,mrv_B_2014B, mrv_B_2014J)

#there are significant amount of rows lost during the merging process put it down to the fact that there were data missing in  vle file 
#otherwise possible never entered students to vle where no clicks - unregistered - check 
vle_B%>%
  filter(sum_click == 0)
#there is no 0 count vle available means that students who had no clicks on website were never documented in this file 
#check unregistered students before start date 
reg_B %>%
  filter(date_unregistration <= 0)
#there are 988 student unregistered before module B started 
#there is still some row loss is unaccounted (299) but the majority of data loss is explainable

```
```{r}
#CCC
main_C <- main%>%
  filter(code_module == "CCC")
unique(main_C$code_presentation)
#2014B
main_C_2014B <- main_C%>%
  filter(code_presentation == "2014B")
#2014J
main_C_2014J <- main_C %>%
  filter(code_presentation == "2014J")

#reg
reg_C <- reg%>%
  filter(code_module == "CCC")
reg_C_2014B <- reg_C%>%
  filter(code_presentation == "2014B")%>%
   select(id_student,date_unregistration)
reg_C_2014J <- reg_C %>%
  filter(code_presentation == "2014J")%>%
   select(id_student,date_unregistration)

#vle
vle_C <- vle %>%
  filter(code_module == "CCC")
vle_C_2014B <- vle_C %>%
  filter(code_presentation == "2014B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_C_2014J <- vle_C %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merge
mr_C_2014B <- merge(main_C_2014B,reg_C_2014B,by="id_student")
#add vle
mrv_C_2014B <- merge(mr_C_2014B, vle_C_2014B, by="id_student")
mr_C_2014J <- merge(main_C_2014J,reg_C_2014J,by="id_student")
#add vle
mrv_C_2014J <- merge(mr_C_2014J, vle_C_2014J, by="id_student")

CCC <- rbind(mrv_C_2014B,mrv_C_2014J)
```

```{r}
#DDD
main_D <- main %>%
  filter(code_module == "DDD")
unique(main_D$code_presentation)

#there are four different files needed 
main_D_2013B <- main_D%>%
  filter(code_presentation == "2013B")
main_D_2013J <- main_D %>%
  filter(code_presentation =="2013J")
main_D_2014B <- main_D %>%
  filter(code_presentation == "2014B")
main_D_2014J <- main_D %>%
  filter(code_presentation == "2014J")

#break registration
reg_D <- reg %>%
  filter(code_module == "DDD")
reg_D_2013B <- reg_D%>%
  filter(code_presentation == "2013B")%>%
    select(id_student,date_unregistration)
reg_D_2013J <- reg_D %>%
  filter(code_presentation =="2013J")%>%
    select(id_student,date_unregistration)
reg_D_2014B <- reg_D %>%
  filter(code_presentation == "2014B")%>%
    select(id_student,date_unregistration)
reg_D_2014J <- reg_D %>%
  filter(code_presentation == "2014J")%>%
    select(id_student,date_unregistration)

#vle
vle_D <- vle %>%
  filter(code_module == "DDD")
vle_D_2013B <- vle_D %>%
  filter(code_presentation == "2013B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_D_2013J <- vle_D %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_D_2014B <- vle_D %>%
  filter(code_presentation == "2014B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_D_2014J <- vle_D %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merging 

mr_D_2013B <- merge(main_D_2013B,reg_D_2013B,by="id_student")
#add vle
mrv_D_2013B <- merge(mr_D_2013B, vle_D_2013B, by="id_student")
#D_2013J
mr_D_2013J <- merge(main_D_2013J,reg_D_2013J,by="id_student")
#add vle
mrv_D_2013J <- merge(mr_D_2013J, vle_D_2013J, by="id_student")
#2014_B
mr_D_2014B <- merge(main_D_2014B,reg_D_2014B,by="id_student")
#add vle
mrv_D_2014B <- merge(mr_D_2014B, vle_D_2014B, by="id_student")
#D_2014J
mr_D_2014J <- merge(main_D_2014J,reg_D_2014J,by="id_student")
#add vle
mrv_D_2014J <- merge(mr_D_2014J, vle_D_2014J, by="id_student")

DDD <- rbind(mrv_D_2013B,mrv_D_2013J,mrv_D_2014B, mrv_D_2014J)

```
```{r}
#EEE
main_E <- main %>%
  filter(code_module == "EEE")
unique(main_E$code_presentation)

#there are three different files needed 
main_E_2013J <- main_E %>%
  filter(code_presentation =="2013J")
main_E_2014B <- main_E %>%
  filter(code_presentation == "2014B")
main_E_2014J <- main_E %>%
  filter(code_presentation == "2014J")

#break registration
reg_E <- reg %>%
  filter(code_module == "EEE")
reg_E_2013J <- reg_E %>%
  filter(code_presentation =="2013J")%>%
    select(id_student,date_unregistration)
reg_E_2014B <- reg_E %>%
  filter(code_presentation == "2014B")%>%
    select(id_student,date_unregistration)
reg_E_2014J <- reg_E %>%
  filter(code_presentation == "2014J")%>%
    select(id_student,date_unregistration)

#vle
vle_E <- vle %>%
  filter(code_module == "EEE")
vle_E_2013J <- vle_E %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_E_2014B <- vle_E %>%
  filter(code_presentation == "2014B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_E_2014J <- vle_E %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merging 
#E_2013J
mr_E_2013J <- merge(main_E_2013J,reg_E_2013J,by="id_student")
#add vle
mrv_E_2013J <- merge(mr_E_2013J, vle_E_2013J, by="id_student")
#2014_B
mr_E_2014B <- merge(main_E_2014B,reg_E_2014B,by="id_student")
#add vle
mrv_E_2014B <- merge(mr_E_2014B, vle_E_2014B, by="id_student")
#E_2014J
mr_E_2014J <- merge(main_E_2014J,reg_E_2014J,by="id_student")
#add vle
mrv_E_2014J <- merge(mr_E_2014J, vle_E_2014J, by="id_student")

EEE <- rbind(mrv_E_2013J,mrv_E_2014B, mrv_E_2014J)

```

```{r}
#FFF
main_F <- main %>%
  filter(code_module == "FFF")
unique(main_F$code_presentation)

#there are four different files needed 
main_F_2013B <- main_F%>%
  filter(code_presentation == "2013B")
main_F_2013J <- main_F %>%
  filter(code_presentation =="2013J")
main_F_2014B <- main_F %>%
  filter(code_presentation == "2014B")
main_F_2014J <- main_F %>%
  filter(code_presentation == "2014J")

#break registration
reg_F <- reg %>%
  filter(code_module == "FFF")
reg_F_2013B <- reg_F%>%
  filter(code_presentation == "2013B")%>%
    select(id_student,date_unregistration)
reg_F_2013J <- reg_F %>%
  filter(code_presentation =="2013J")%>%
    select(id_student,date_unregistration)
reg_F_2014B <- reg_F %>%
  filter(code_presentation == "2014B")%>%
    select(id_student,date_unregistration)
reg_F_2014J <- reg_F %>%
  filter(code_presentation == "2014J")%>%
    select(id_student,date_unregistration)

#vle
vle_F <- vle %>%
  filter(code_module == "FFF")
vle_F_2013B <- vle_F %>%
  filter(code_presentation == "2013B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_F_2013J <- vle_F %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_F_2014B <- vle_F %>%
  filter(code_presentation == "2014B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_F_2014J <- vle_F %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merging 
mr_F_2013B <- merge(main_F_2013B,reg_F_2013B,by="id_student")
#add vle
mrv_F_2013B <- merge(mr_F_2013B, vle_F_2013B, by="id_student")
#F_2013J
mr_F_2013J <- merge(main_F_2013J,reg_F_2013J,by="id_student")
#add vle
mrv_F_2013J <- merge(mr_F_2013J, vle_F_2013J, by="id_student")
#2014_B
mr_F_2014B <- merge(main_F_2014B,reg_F_2014B,by="id_student")
#add vle
mrv_F_2014B <- merge(mr_F_2014B, vle_F_2014B, by="id_student")
#F_2014J
mr_F_2014J <- merge(main_F_2014J,reg_F_2014J,by="id_student")
#add vle
mrv_F_2014J <- merge(mr_F_2014J, vle_F_2014J, by="id_student")

FFF <- rbind(mrv_F_2013B,mrv_F_2013J,mrv_F_2014B, mrv_F_2014J)
```


```{r}
#GGG
main_G <- main %>%
  filter(code_module == "GGG")
unique(main_G$code_presentation)

#there are three different files needed 
main_G_2013J <- main_G %>%
  filter(code_presentation =="2013J")
main_G_2014B <- main_G %>%
  filter(code_presentation == "2014B")
main_G_2014J <- main_G %>%
  filter(code_presentation == "2014J")

#break registration
reg_G <- reg %>%
  filter(code_module == "GGG")
reg_G_2013J <- reg_G %>%
  filter(code_presentation =="2013J")%>%
    select(id_student,date_unregistration)
reg_G_2014B <- reg_G %>%
  filter(code_presentation == "2014B")%>%
    select(id_student,date_unregistration)
reg_G_2014J <- reg_G %>%
  filter(code_presentation == "2014J")%>%
    select(id_student,date_unregistration)

#vle
vle_G <- vle %>%
  filter(code_module == "GGG")
vle_G_2013J <- vle_G %>%
  filter(code_presentation == "2013J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_G_2014B <- vle_G %>%
  filter(code_presentation == "2014B")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))
vle_G_2014J <- vle_G %>%
  filter(code_presentation == "2014J")%>%
  select(id_student,sum_click)%>%
  group_by(id_student)%>%
  summarise(clicks = sum(sum_click))

#merging 
mr_G_2013J <- merge(main_G_2013J,reg_G_2013J,by="id_student")
#add vle
mrv_G_2013J <- merge(mr_G_2013J, vle_G_2013J, by="id_student")
#2014_B
mr_G_2014B <- merge(main_G_2014B,reg_G_2014B,by="id_student")
#add vle
mrv_G_2014B <- merge(mr_G_2014B, vle_G_2014B, by="id_student")
#G_2014J
mr_G_2014J <- merge(main_G_2014J,reg_G_2014J,by="id_student")
#add vle
mrv_G_2014J <- merge(mr_G_2014J, vle_G_2014J, by="id_student")

GGG <- rbind(mrv_G_2013J,mrv_G_2014B, mrv_G_2014J)

```





All separate merging carried out the individual files can be added back together to create the main working file. 

```{r}
main_data <- rbind(AAA,BBB,CCC,DDD,EEE,FFF,GGG)

```

There are `r 32593-29228` row lost in the process identified as those students whom unregistered from the course before its started and had 0 clicks on the module website. 

Write out the data file for EDA and further work. Proceed to next stage #'Part_2'. 

```{r}
#write out file
write.csv(main_data,"~/git/OU Learning Analytics/Raw Data/main_data.csv")
```



